---
title: "Main Analysis"
author: "Jonas Han (jh3877)"
date: "December 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(reshape2)
```


# Introduction

### Team
- Jonas Han (jh3877): Introduction, Executive Summary, data cleaning, GitHub folder/file structure
- Justin Ali Kennedy (jak2294):
- Lerong Chen (lc3375):
- Sidney Frank Fletcher (sff2114):

Stop-and-Frisk is a NYPD practice of briefly detaining, questioning, and sometimes frisking or searching civilians for weapons and/or contraband. It was quite prevalent during then-Mayor Michael Bloomberg's tenure as he was a strong proponent of the program. Unsuprisingly, stop-and-frisk became quite controversial during the height of the program in 2011. Many critics of the program claimed that the NYPD was using racial profiling when selecting suspects, specifically Black and Hispanic suspects. Other critisized the program's efficacy in reducing the crime rate in NYC. Proponents of the practice argued that stop-and-frisk made the city safer, especially in neighborhoods with high Black and Hispanic populations.

Our motivation in selecting this project is to analyze the stop-and-frisk data and determine for ourselves if the NYPD was indeed using racial profiling to select suspects. Additionally, we wanted to understand the program's effectiveness in actually deterring and/or stopping crime in NYC. Lastly, one of our team members (Jonas) was a high school student in Brooklyn around the peak of the stop-and-frisk program and witnessed many stops in the train station nearby his school. 


# Executive Summary

From the frequency bar plot below, we observe that the practice of stop-and-frisk increased significantly during then-Mayor Michael Bloomberg's tenure in office, which began in 2002 and ended in 2013. The practice peaked in 2011, with over 650,000 stops, before being severely halted in 2013 and 2014. The subsequent sharp decline in stop-and-frisk was led by a federal lawsuit against the practice in 2013 as well as Bill de Blasio's election to the mayor's office in 2014. Bill de Blasio was an avid opponent of stop-and-frisk and ran a campaign promising to reform the practice. Since taking office in 2014, stop-and-frisk has continued to decrease every single year. 

```{r fig.height=7, fig.width=12}
sqf_by_year <- read.csv(file="./Data/sqf_by_year.csv", header=TRUE)
ggplot(data=sqf_by_year, aes(x=year, y=count/1000)) + geom_bar(stat="identity") + ggtitle("NYPD Stop-and-Frisk (2003-2017)") + theme_grey(18) + theme(plot.title=element_text(hjust=0.5)) + scale_x_continuous("Year", labels=as.character(sqf_by_year$year), breaks=sqf_by_year$year) + ylab("Frequency (Thousands)")
```


For an preliminary analysis, we compared the race of stop-and-frisk suspects against the demographics of NYC from the 2010 Census. From this initial perspective, it appears that there may be an issue of racial profiling in regards to the stop-and-frisk program. From 2003 to 2017, over 50% of stop-and-frisk suspects were Black while less than 25% of NYC is Black. Furthermore, White suspects accounted for less than 10% of stops, but represent the majority(>30%) of NYC residents. A similar disparity exists for Asians across the two datasets. Note that it was difficult to compare the Hispanic populations across the two different data sources as the stop-and-frisk dataset buckets all Hispanics into White Hispanics or Black Hispanics. In the 2010 Census data, most Hispanics responded in the "Some Other Race alone" category and not in the "White alone" or "Black or African-American alone" categories. 


```{r fig.height=7, fig.width=12}
arrests_by_race <- read.csv(file="./Data/arrests_by_race.csv", header=TRUE)
census_by_race <- read.csv(file="./Data/census_by_race.csv", header=TRUE)
complaints_by_susp_race <- read.csv(file="./Data/complaints_by_susp_race.csv", header=TRUE)
shootings_by_perp_race <- read.csv(file="./Data/shootings_by_perp_race.csv", header=TRUE)
sqf_by_race <- read.csv(file="./Data/sqf_by_race.csv", header=TRUE)

race_data <- do.call("rbind", list(arrests_by_race, complaints_by_susp_race, shootings_by_perp_race, sqf_by_race))
race_data$dataset <- factor(race_data$dataset, levels=c("Stop-and-Frisk", "Complaints", "Arrests", "Shootings"))

par(mfrow=c(1,2))
par(mar = c(11,4,3,1) + 0.1)

barplot(sqf_by_race$percentage, names.arg=sqf_by_race$race, main="Stop-and-Frisk by Suspect Race", ylim=c(0, .55), las=2, ylab="Percentage", yaxt="n")
axis(2, at=pretty(sqf_by_race$percentage), lab=paste0(pretty(sqf_by_race$percentage) * 100, "%"), las=2)

barplot(census_by_race$percentage, names.arg=census_by_race$race, main="NYC 2010 Census by Race", ylim=c(0, .55), las=2, ylab="Percentage", yaxt="n")
axis(2, at=pretty(sqf_by_race$percentage), lab=paste0(pretty(sqf_by_race$percentage) * 100, "%"), las=2)
```


In response to the racial profiling controversy, then-Mayor Michael Bloomberg stated that "If we stopped people based on census numbers, we would stop many fewer criminals, recover many fewer weapons and allow many more violent crimes to take place". He has also further stated in an op-ed for The Washington Post that "Ninety percent of all people killed in our city — and 90 percent of all those who commit the murders and other violent crimes — are black and Hispanic". We were interested to find out for ourselves if the former mayor's statements were true. Thus, we looked at the suspect/perpetrator race variable across the stop-and-frisk, complaints, arrests, and shooting incidents datasets to answer the question. From the plot below, we observe that the racial composition of the suspects/perpetrators in stop-and-frisk, complaints, and arrests datasets were all fairly similar. That is, it appears that Blacks are stopped disproportionately compared to the percentage of Blacks in NYC, but not compared to the percentage of complaints filed against Black suspects or arrests of Black perpetrators. Shooting incidents in NYC skew even more heavily towards Black perpetrators. This supoprts Michael Bloomberg's claim that "the proportion of stops generally reflects our crime numbers". There are two possible explanations for this. The first is that society is prejudiced against Blacks and thus are more likely to file complaints against Blacks or claim that a Black suspect committed a crime. Another possibility is that Blacks commit crime at disproportionate rates compared to other races. 


```{r fig.height=5, fig.width=12}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

race_bar_chart <- ggplot(race_data, aes(x=race, y=percentage, fill=race)) + geom_bar(stat="identity") + facet_grid(.~dataset) + theme(axis.text.x = element_text(angle = 60, hjust=1)) + xlab("Suspect/Perpetrator Race") + ylab("Percentage") + guides(fill=guide_legend(title="")) + scale_fill_manual(values=cbbPalette) + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
race_bar_chart
```

Next, we compared the suspect/perpetrator race against the victim race for the complaints and shooting incidents datasets. We wanted to discover how closely the victim race distributions aligned with the suspect/perpetrator race distributions. Across all four bar plots below, Blacks are overrepresented compared to the NYC Census data. That is, Blacks disproportionantely commit more crime, but are also more likely to be victims of crime. It is also interesting to note that complaint victims skew less towards Blacks and more towards Whites and Asians compared to complaint suspects. This is in contrast to shooting incidents, where the distribution of perpetrator race is nearly identical to the distribution of victim race. This may indicate that shootings typically involve perpetrators and victims of the same race. This is a topic that we would have liked to researched further, given more time.

```{r fig.height=5, fig.width=12}
complaints_by_vic_race <- read.csv(file="./Data/complaints_by_vic_race.csv", header=TRUE)
complaints_by_vic_race$susp_vic_category <- "Complaints | Victim"
shootings_by_vic_race <- read.csv(file="./Data/shootings_by_vic_race.csv", header=TRUE)
shootings_by_vic_race$susp_vic_category <- "Shootings | Victim"

complaints_by_susp_race$susp_vic_category <- "Complaints | Suspect"
shootings_by_perp_race$susp_vic_category <- "Shootings | Perpetrator"

susp_vic_by_race <- do.call("rbind", list(complaints_by_vic_race, complaints_by_susp_race, shootings_by_perp_race, shootings_by_vic_race))
susp_vic_by_race$susp_vic_category <- factor(susp_vic_by_race$susp_vic_category, levels=c("Complaints | Suspect", "Complaints | Victim", "Shootings | Perpetrator", "Shootings | Victim"))


cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

susp_vic_bar_chart <- ggplot(susp_vic_by_race, aes(x=race, y=percentage, fill=race)) + geom_bar(stat="identity") + facet_grid(.~susp_vic_category) + theme(axis.text.x = element_text(angle = 60, hjust=1)) + xlab("") + ylab("Percentage") + guides(fill=guide_legend(title="")) + scale_fill_manual(values=cbbPalette) + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
susp_vic_bar_chart
```

The vast majority of stops(approx. 88%) involved innocent suspects and only 12% of stops led to an arrest or summons. Furthermore, only about 3% of stops found a weapon or contraband. It is difficult for us to judge whether or not these catch rates are effective as we do not have another metric to compare it to. Critics of stop-and-frisk may argue that this may be a wasteful use of police resources since the vast majority of stops involve innocent civilians. However, proponents of the practice may argue that over the course of the last 15 years (2003-2017), an extra 50,000 weapons have been taken off the streets, 100,000 contraband were confiscated, and 300,000 criminals were arrested. 

```{r fig.height=5, fig.width=9}
not_innocent_df <- read.csv(file="./Data/sqf_not_innocent.csv", header=TRUE)

ggplot(not_innocent_df, aes(x=reorder(Category, Count), y=Count/1000)) + geom_bar(stat="identity") + theme_grey(16) + coord_flip() + ylab("Frequency(Thousands)") + xlab("") + theme(plot.margin = margin(10, 15, 10, 1)) + ggtitle("Stop-and-Frisk Efficacy")

ggplot(not_innocent_df, aes(x=reorder(Category, Percentage_of_Stops), y=Percentage_of_Stops)) + geom_bar(stat="identity") + theme_grey(16) + coord_flip() + ylab("Percentage of Stops") + xlab("") + scale_y_continuous(labels = scales::percent) + theme(plot.margin = margin(10, 15, 10, 1)) + ggtitle("Stop-and-Frisk Efficacy")
```

We will next analyze the effect that stop-and-frisk had on crime levels in NYC. Specifically, we will compare the levels of stop-and-frisk against the levels of crime for both misdemeanors and felonies from 2003 to 2017. The NYC crime statistics page classifies felonies into two categories: "Citywide Seven Major Felony Offenses" and "Citywide Non-Seven Major Felony Offenses". The seven major felonies inlcude crimes such as murder, rape, robbery, felony assault, burglary, grand larceny, and grand larceny of a motor vehicle. The non-seven major felonies include arson, drug possession, weapons possession, fraud, forgery, etc. Since the four categories of data (stop-and-frisk, major felonies, non-major felonies, misdemeanors) that we wanted to plot together have vastly different scales, we indexed all the data points according to their respective starting values in 2003. As evinced from the line plot below, all types of crime have steadily decreased from 2003 to 2017. Stop-and-frisk, on the other hand, increased quickly throughout the mid to late 2000s and peaked in 2011. We observe that during this same time period, non-major felonies and misdemeanors have remained steady while major felonies did experience a noticeable decline. One could argue that this could have been due to the increased instances of stop-and-frisk throughout NYC. However, after the election of Bill de Blasio and the subsequent overhaul of the stop-and-frisk program, we notice that crimes did not increase at all. In fact, crime for all three categories (major felonies, non-major felonies, misdemeanors) have continued to decline. Moreover, 2017 has seen the lowest level of crime in all the years we analyzed while simultaneously also having the least amount of stop-and-frisks. This indicates that stop-and-frisk may not have had much of an impact on crime rate in NYC as otherwise suggested by Michael Bloomberg.

```{r fig.height=7, fig.width=12}
# https://www1.nyc.gov/site/nypd/stats/crime-statistics/historical.page
sqf_vs_crime_by_year <- sqf_by_year
colnames(sqf_vs_crime_by_year) <- c("year", "sqf_count")
major_felonies_by_year <- c(147069, 142093,	135475,	128682,	121009,	117956,	106730,	105115,	106669,	111147,	111335,	106722,	105453, 101716, 96658)
non_major_felonies_by_year <- c(61217, 66733, 67854, 69028, 70958, 68958, 63760, 59387, 57240, 56902, 57650, 56869, 56520, 58346, 54907)
misdemeanors_by_year <- c(365471, 357724, 353649, 361574, 378616, 380406, 386765, 391892, 383108, 374364, 359350, 348371, 322848, 314925, 300354)

sqf_vs_crime_by_year$major_felonies <- major_felonies_by_year
sqf_vs_crime_by_year$nonmajor_felonies <- non_major_felonies_by_year
sqf_vs_crime_by_year$misdemeanors <- misdemeanors_by_year

starting_values_df <- sqf_vs_crime_by_year[sqf_vs_crime_by_year$year == 2003,]
sqf_vs_crime_by_year["Stop-and-Frisk"] <- (sqf_vs_crime_by_year$sqf_count/starting_values_df$sqf_count) * 100
sqf_vs_crime_by_year["Major Felonies"] <- (sqf_vs_crime_by_year$major_felonies/starting_values_df$major_felonies) * 100
sqf_vs_crime_by_year["Non-Major Felonies"] <- (sqf_vs_crime_by_year$nonmajor_felonies/starting_values_df$nonmajor_felonies) * 100
sqf_vs_crime_by_year["Misdemeanors"] <- (sqf_vs_crime_by_year$misdemeanors/starting_values_df$misdemeanors) * 100

sqf_crime_indexed_by_year <- melt(sqf_vs_crime_by_year, id.vars = "year", measure.vars = c("Stop-and-Frisk", "Major Felonies", "Non-Major Felonies", "Misdemeanors"))

colnames(sqf_crime_indexed_by_year)[2] <- "Category"

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot() + geom_line(data=sqf_crime_indexed_by_year, aes(x=year, y=value, color=Category), lwd=1.5) + xlab("Year") + ylab("Index") + ggtitle("Stop-and-Frisk vs. Crime") + scale_color_manual(values=cbbPalette) + theme_grey(16) + theme(plot.title=element_text(hjust=0.5)) + scale_x_continuous("", labels = as.character(sqf_crime_indexed_by_year$year), breaks = sqf_crime_indexed_by_year$year)

```

```{r fig.height=6, fig.width=11}
sqf_race_agg_by_year <- read.csv("./Data/sqf_race_by_year.csv", header=TRUE)

ggplot() + geom_line(data=sqf_race_agg_by_year, aes(x=year, y=percentage, color=race), lwd=1.5) + ylab("Percentage") + ggtitle("Stops by Race Across Years") + guides(fill=guide_legend(title="")) + scale_color_brewer(palette="Set1") + theme_grey(16) + theme(plot.title=element_text(hjust=0.5)) + scale_x_continuous("", labels = as.character(sqf_race_agg_by_year$year), breaks = sqf_race_agg_by_year$year) + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```























